### 2017-06-03 tex/paper_v1.tex
init tex

%% LyX 2.2.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,sigconf]{acmart}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{color}
\usepackage{amsmath}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\usepackage{booktabs}% For formal tables
\usepackage{float}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\usepackage{rotating}
\usepackage{array}
\newcolumntype{L}[1]{>{\vspace{0.5em}\begin{minipage}{#1}\raggedright\let\newline\\
\arraybackslash\hspace{0pt}}m{#1}<{\end{minipage}\vspace{0.5em}}}
\newcolumntype{R}[1]{>{\vspace{0.5em}\begin{minipage}{#1}\raggedleft\let\newline\\
\arraybackslash\hspace{0pt}}m{#1}<{\end{minipage}\vspace{0.5em}}}
\newcolumntype{C}[1]{>{\vspace{0.5em}\begin{minipage}{#1}\centering\let\newline\\
\arraybackslash\hspace{0pt}}m{#1}<{\end{minipage}\vspace{0.5em}}}

\copyrightyear{2017}
\acmYear{2017}
\setcopyright{acmcopyright}
\acmConference{}{}{} 
\acmPrice{15.00} 
\acmDOI{http://dx.doi.org/xxx}
\acmISBN{ACM ISBN xxx}

\clubpenalty=10000 
\widowpenalty = 10000

\makeatother

\usepackage{babel}
\begin{document}
\title{Frame-Transformer Emotion Classification Network}

\author{Jiarui~Gao$^1$}

\affiliation{
\institution{  $^1$School of Computer Science, Shanghai Key Laboratory of Intelligent Information Processing,\\ Fudan University, China}   
}
\email{  jrgao14@fudan.edu.cn}

\begin{abstract}

Emotional content is a key ingredient in user-generated videos.  
However, due to the emotions sparsely expressed in the user-generated video,  it is very difficult to analayze emotions in  videos. In this paper, we propose a new architecture--Frame-Transformer Emotion Classification Network (FT-EC-net) to solve three highly correlated emotion analysis tasks:  emotion recognition, emotion attribution and emotion-oriented summarization. We also contribute a new dataset for emotion attribution task by annotating the ground-truth labels of attribution segments. A comprehensive set of experiments on two datasets demonstrate the effectiveness of our framework.
\end{abstract}
\keywords{Video emotion recognition; video emotion attribution; video emotion-oriented summarization and spatial-transformer network}

\maketitle

\bibliographystyle{ACM-Reference-Format}
\bibliography{security} 
\end{document}